---
title     : 'R, FRED, and the 2016 Texas Primary: Part 2'
author    : "Luke Smith"
date      : "April 24, 2016"
framework : io2012
---

## Overview
#### Picking up from last time

I have all the metadata and the data itself. However, all that data is floating around in data frames inside a list. I need a way to clean this up to make operations on the data frames more manageable.

---

## `cat.functions.R`
#### Introduction

The file `cat.functions.R` contains a few functions that will help us tidy up this data:

* `cat.info()` = creates a master/top-level table
* `obs.catcher()` = uses 'category' information from cat.info() to group data frames in fred.obs by 'category'
* `cat.tabler()` = creates tables each 'category' list element returned by obs.catcher()

---

## `cat.functions.R`
#### Why create and use `cat.functions.R`?

1. The first function, `cat.info()`, will create a handy table with information about each category
2. We can use this table to check the frequency and range of each category
  a. Some categories will require aggregating monthly data into annual data
  b. Some categories will limit our ability to use the most recent data in other categories
3. The second function, `obs.catcher()`, will categorize the data in `fred.obs`
4. The last function in the list, `cat.tabler()`,

---

```{r, echo = F, results = F, }
```
## `cat.functions.R`
#### `cat.info()`

This will create a master/top-level table from which to draw certain metadata which will be used in other functions.

```{r, results = F, eval = F}
cat.info <- function(series.table){
    
### load dependencies
    require(plyr, quietly = T) ## I have a habit of loading plyr BEFORE dplyr
    require(dply, quietly = T)

### make an aggregate table based on $Category
    summarized.series <- fred.series %>% select(2,3,6,7,8,9) %>%
                         aggregate(list(series.table$Category), unique) %>% select(3,2,4:7) %>%
                         arrange(Release, Category)
}
```

---

## `cat.functions.R`
#### `obs.catcher()`

This will categorize the data frames in the `fred.obs` list. With this categorized list, we will be able to aggregate data as needed for our analysis (i.e. summarizing monthly data into annual data, etc).

```{r, results = F, eval = F}
obs.catcher <- function(series.table, obs.list, cat){
    
### subset by category
    cat.series               <- series.table[series.table$Category == cat, ]
    
### Find all SeriesID matching the subset's SeriesID
    cat.obs <- list()
    cat.obs <- obs.list[cat.series$SeriesID]
}
```

---

## `cat.functions.R`
#### `cat.tabler()`

This will create a table from each category in the `fred.obs` list. The first column is `Date` and has a further 254 columns (one for each county). The 254 columns have column headers that correspond to the `SeriesID`.

```{r, results = F, eval = F, }
cat.tabler <- function(obs.list){
    x <- 1
    while(x <= length(obs.list)){
        if(x == 1){
            main.frame           <- data.frame() # initialize the data frame
            main.frame           <- obs.list[[x]] # add the first list object
            names(main.frame)[2] <- names(obs.list)[1] # give that object a name
            main.frame
        }   else {
            main.frame               <- merge(main.frame, obs.list[[x]], by = "Date")
            names(main.frame)[x + 1] <- names(obs.list)[x]
            main.frame
        }
        x = x + 1
        main.frame
    }
    main.frame
}
```

---

## Running the functions
#### Plan of attack

Here is our plan of attack for using these functions to tidy our data:

1. Create a top-level/master table using `cat.info()`
2. Use the result table from `cat.info()` to group the data frames in `fred.obs` by category
  a. We will use `obs.catcher()` to do this
3. Create separate tables of each category
  a. Each table will correspond to one of the categories in `cat.info()`
  b. Each table will have 255 columns (one called `Date` and the others will have names corresponding to the `SeriesID` of each of the 254 counties in Texas)

---

## Running the functions
#### Loading the dependencies

```{r, results = F, eval = F}
library(plyr, quietly = T)
library(dplyr, quietly = T)
library(rvest, quietly = T)
# library(choroplethrMaps, quietly = T)
library(lubridate, quietly = T)
source("cat.functions.R")
load("Data/fred.series")
load("Data/fred.obs")
```

Easy so far.

---

## Running the functions
#### Running `cat.info()`, `obs.catcher()`, and `cat.tabler()`

* Create a top-level/master table using `cat.info()`

```{r, results = F, eval = F}
fred.master <- cat.info(fred.series)
```

* Group the data frames in `fred.obs` by category using `obs.catcher()`

```{r, results = F, eval = F}
 fred.cat.list <- lapply(seq_along(fred.master$category), function(x){
        obs.catcher(series.table = fred.series,
                    obs.list     = fred.obs,
                    cat          = fred.master$category[x])
    })
names(fred.cat.list) <- fred.master$category
save(fred.cat.list, file = "Data/fred.cat.list.RData")
```

* Create separate tables of each category using `cat.tabler()`

```{r, results = F, eval = F}
 fred.tables <- lapply(seq_along(fred.cat.list), function(x){
        cat.tabler(fred.cat.list[[x]])
    })
names(fred.tables) <- fred.master$category
# fred.tables   <- fred.tables %>% select(-1) %>% as.character() %>% as.numeric()
save(fred.tables, file = "Data/fred.tables.RData")
```

---

## Results

Let's take a look at what we have.

```{r, echo = F, results = F}
load("~/R/TexasPrimary2016/Data/fred.tables.RData")
load("~/R/TexasPrimary2016/Data/fred.master.RData")
load("~/R/TexasPrimary2016/Data/fred.cat.list.RData")
```

* `cat.info()`
```{r}
head(fred.master)
```

* `obs.catcher()`
```{r}
tail(fred.cat.list$`Resident Population`$TXHARR1POP)
```

* `cat.tabler()`
```{r}
head(fred.tables[[1]][1:5])
```